---
title: "Assignment Practical Machine learning"
author: "Joris Meulenbeek"
date: "26 augustus 2016"
output: html_document
---

#Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 
```{r}

```

#Data 

The training data for this project are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 

#What you should submit

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

#Loading libraries
For this assigment I used the following packages:


```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
```

#Loading data files
I manually downloaded the .csv files from:
The training data for this project are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

I saved the files in the directory data. 

Now I gonna set my working directory:
```{r}
setwd("C:/Users/Joris/Documents/Practical machine learning/Week 4")
```

And then load the files in to R, where the values "#DIV/0! and empty are also NA (they have to go out later). 
```{r}
In_train <- read.csv("./data/pml-training.csv", na.strings=c("NA","#DIV/0!",""))
```
```{r}
In_validate <- read.csv("./data/pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
```

#Data exploration
I want to see which data is in the Coursera file for training and validation.
```{r}
dim(In_train)
str(In_train)
```
```{r}
dim(In_validate)
str(In_validate)
```

#Remving columns with NA
Now we gonna remove all collumns with only NA in the validtion set. Also the first 7 columns will be removed cus they are not important for the model. 
```{r}
Right_col <- names(In_validate[,colSums(is.na(In_validate)) == 0])[8:59]
```
```{r}
In_train <- In_train[,c(Right_col,"classe")]
```
```{r}
In_validate <- In_validate[,c(Right_col,"problem_id")]
```
```{r}
dim(In_train); dim(In_validate);
```
You can see that there are only 53 columns (from the original 160) are still there.

#Making train and test set
As we been teached, the train set has to be around 60% of the data and the test set around 40%. 

```{r}
set.seed(12345)
inTrain <- createDataPartition(In_train$classe, p=0.6, list=FALSE)
```

```{r}
training <- In_train[inTrain,]
testing <- In_train[-inTrain,]
```

Here you see the number of cases in the train set and test set.
```{r}
dim(training); 
dim(testing);
```
#Decision Tree Model
The first model will be a decision tree. The accuracy will not be very high.
But we will see.

```{r}
modFit <- rpart(classe ~ ., data = training, method="class")
fancyRpartPlot(modFit)
```
#Predicting with Decision Tree
```{r}
set.seed(12345)
prediction <- predict(modFit, testing, type = "class")
confusionMatrix(prediction, testing$classe)
```
The accuracy is around 73% (and that low)

#Random Forest Model
The second model will be a Random Forrest. The accuracy will be way higher then the tree.
```{r}
set.seed(12345)
modFitRF <- randomForest(classe ~ ., data = training)
```
#Predicting with Random Forrest
```{r}
prediction <- predict(modFitRF, testing, type = "class")
confusionMatrix(prediction, testing$classe)
```
The accuracy is around 99,5% (and that very high)

##Predicting on the Validation Data
Now we gonna see what both models will do on the validation set.

Decision Tree Prediction
```{r}
prediction_tree <- predict(modFit, In_validate, type = "class")
prediction_tree
```
Random Forest Prediction
```{r}
predictionRF <- predict(modFitRF, In_validate, type = "class")
predictionRF
```
The random forrest model is very accurate. Thats why all of the test cases will be predict okay for 99%. 
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
```
pml_write_files(predictionRF)
